Mixed language processing increases cross-language phonetic transfer in Bengali-English bilinguals

### 2022-10-11 draft_BLC/blc_revision1_reviewer_response.txt

The third comment of substance, which I hadn’t noticed before, has to do with acoustic analyses. On page 26, it is said that formant values were Lobanov-normalized. I don’t think this is possible in this study. I believe the authors did something akin to Lobanov-normalization but quite distinct from it. In Lobanov normalization, values are z-scored with reference to values of all the vowels in a speaker’s vowel system. In other words, for one to be able to do a Lobanov-normalization for a given vowel category (the tokens from a category) one MUST have tokens, collected from the same speaker, of all the other vowels in the speaker’s vowel system. In this case, the given vowel token is expressed with reference to its location in a speaker’s vowel space. This DOES allow to normalize across genders. This is a great normalization procedure, but one needs to have collected samples of all the vowels in the system, or at least of the corner vowels, which outline the limits of the system. I don’t think this is what the authors did here (because I don’t think they collected samples of all the vowels of English or at least the corner vowels). Their normalization method is akin to the Lobanov method in that they did z-scoring with all the vowels they had. However, if they didn’t have samples of all the vowels, this is not a true normalization method, and clearly not Lobanov’s normalization procedure. Please, explain better and then explain what this normalization method accomplished. (Since mixed-effects regression was to be used, and all factors are within-speaker, normalization was likely unnecessary.)

AR: Thanks for pointing this out; we agree that "Lobanov normalization" is not the appropriate term here, given the lack of all corner vowels. We have replaced this with the term "z-score normalization". The target was to represent all formant values on a uniform scale across speakers, removing speaker-specific value ranges, so that the relative positions of the two vowels and shift between conditions could be compared directly. Since the data had productions of the same two categories from all speakers, we felt comfortable that this would not affect the interpretability of the values. This was useful for first-pass visualizations to understand speakers' vowel systems (e.g. fig. 2 in the text, which shows that speakers vary greatly in their default realizations of the two IE vowels), and to visualize gender differences without confounds from individual differences (e.g. figure 6). Since we were expecting high individual variability, we did this before running any further analyses. We agree that given the specific statistical tools used here, this was not essential to the statistical analyses. We re-ran the analysis with non-normalized values to verify this, and found identical results. We have included the code and output from both analyses in the OSF repository for the project, and added a footnote in the main text addressing this (pg. ...).

### 2022-10-11 draft_BLC/BLC_AM-ID_manuscript.tex
bengali picture task figure

\end{enumerate}
\subsection{Acoustic analysis}\label{analysis}
The first iteration of each task was treated as a trial, and excluded from further analyses. An additional 67 tokens (2.79\%) were excluded due to errors in presentation/production, giving 2333 target word tokens (\nt{2}: 1137, \nt{\ae}: 1196) for analysis. Recordings were manually segmented and annotated in Praat \citep{boersma2016praat}, and the first and second formant frequencies (F1 and F2) of the target vowels \nt{2, \ae} were measured at five points: in 10\% increments starting at 5\% into the vowel. This gave measures of F1 and F2 at 5\%, 15\%, 25\%, 35\% and 45\% into the vowel, allowing us to observe any dynamic changes in the extent of transfer over the course of producing the vowel. Individual vocal tract differences affect formant frequencies across speakers. To reduce between-speaker differences, we used z-score normalization (using the PhonR package \citep{phonR} in R \citep{r}, representing each speaker's formant frequencies as the scaled distance from their own mean). \footnote{Since our experimental treatments were between-subject and statistical analyses use mixed-effects regression, this step would not affect the results, as pointed out by a reviewer. This was done to allow for more intuitive values and visualizations, rater than statistical necessity. The code for parallel statistical analysis using non-z-score-normalized values can be found in the OSF repository.} Plotting these vowels shows a difference between male and female speakers, such that male speakers have overall lower formant frequencies (figure \ref{gender_before}). %Speaker normalization using the z-score method does not get rid of this difference. 
Since gender-related differences in formant frequencies are not of interest in this study, we normalized individual values by gender (representing each value as a scaled distance from the group mean; figure \ref{gender_after}). Statistical analyses were performed on these normalized values.

@@

\begin{figure}[!htb] \label{figure_picture_task}
	\centering
	\begin{subfigure}[t]{0.49\textwidth}
		\centering
		\includestandalone[mode=image]{picture_task_eng}
		\caption{English (nonswitch) trial} \label{picture_task_english}
	\end{subfigure}
	\begin{subfigure}[t]{0.49\textwidth}
		\centering
		\includestandalone[mode=image]{picture_task_bengali} 
		\caption{Bengali-English (switch) trial} \label{picture_trial_bengali}
	\end{subfigure}
	\caption{Picture-naming task sample sequence: language cue (not read out), picture (named as per lang cue), target (read out), distractor}
\end{figure}

>>>>

\end{enumerate}
\subsection{Acoustic analysis}\label{analysis}
The first iteration of each task was treated as a trial, and excluded from further analyses. An additional 67 tokens (2.79\%) were excluded due to errors in presentation/production, giving 2333 target word tokens (\nt{2}: 1137, \nt{\ae}: 1196) for analysis. Recordings were manually segmented and annotated in Praat \citep{boersma2016praat}, and the first and second formant frequencies (F1 and F2) of the target vowels \nt{2, \ae} were measured at five points: in 10\% increments starting at 5\% into the vowel. This gave measures of F1 and F2 at 5\%, 15\%, 25\%, 35\% and 45\% into the vowel, allowing us to observe any dynamic changes in the extent of transfer over the course of producing the vowel. Individual vocal tract differences affect formant frequencies across speakers. To reduce between-speaker differences, we used z-score normalization (using the PhonR package \citep{phonR} in R \citep{r}, representing each speaker's formant frequencies as the scaled distance from their own mean). \footnote{Since our experimental treatments were between-subject and statistical analyses use mixed-effects regression, this step would not affect the results, as pointed out by a reviewer. This was done to allow for more intuitive values and visualizations, rather than statistical necessity. The code and analyses using non-normalized values can be found in the OSF repository.} Plotting these vowels shows a difference between male and female speakers, such that male speakers have overall lower formant frequencies (figure \ref{gender_before}). %Speaker normalization using the z-score method does not get rid of this difference. 
Since gender-related differences in formant frequencies are not of interest in this study, we normalized individual values by gender (representing each value as a scaled distance from the group mean; figure \ref{gender_after}). Statistical analyses were performed on these normalized values.

@@

\begin{figure}[!htb] \label{figure_picture_task}
	\centering
	\begin{subfigure}[t]{0.49\textwidth}
		\centering
		\includegraphics[scale=1]{images/picture_task_eng}
		\caption{English (nonswitch) trial} \label{picture_task_english}
	\end{subfigure}
	\begin{subfigure}[t]{0.49\textwidth}
		\centering
		\includegraphics[scale=1]{images/picture_task_bengali} 
		\caption{Bengali-English (switch/mixed) trial} \label{picture_trial_bengali}
	\end{subfigure}
	\caption{Picture-naming task sample sequence: language cue (not read out), picture (named as per lang cue), target (read out), distractor}
\end{figure}

\begin{comment}
	\begin{figure}[!htb] \label{figure_picture_task}
	\centering
	\begin{subfigure}[t]{0.49\textwidth}
		\centering
		\includestandalone[mode=image]{picture_task_eng}
		\caption{English (nonswitch) trial} \label{picture_task_english}
	\end{subfigure}
	\begin{subfigure}[t]{0.49\textwidth}
		\centering
		\includestandalone[mode=image]{picture_task_bengali} 
		\caption{Bengali-English (switch) trial} \label{picture_trial_bengali}
	\end{subfigure}
	\caption{Picture-naming task sample sequence: language cue (not read out), picture (named as per lang cue), target (read out), distractor}
\end{figure}
\end{comment}
\newpage
